---
title: "Assignment 5 BA 64060"
author: "Samantha Ellis"
date: "2023-11-29"
output: html_document
---

Part One
```{r}
cereal<- na.omit(Cereals) ##Omitted missing values
head(cereal)
Numcereal<- cereal[, unlist(lapply(cereal, is.numeric))]  ##Removed values that were not numeric
NormCereal<- scale(Numcereal[, -1])   ##Scaled/normalized dataset

library(cluster)
hc_singleCereal<- agnes(NormCereal, method = "single")
hc_completeCereal<- agnes(NormCereal, method = "complete")
hc_averageCereal<- agnes(NormCereal, method = "average")
hc_WardCereal<- agnes(NormCereal, method = "ward")

print(hc_singleCereal$ac)
print(hc_completeCereal$ac)
print(hc_averageCereal$ac)
print(hc_WardCereal$ac)
```
The best method for linkage would be the Ward method, because the agglomerative coefficient is the closest to 1, which suggests strong clustering. 




Part Two
```{r}
pltree(hc_WardCereal, cex=0.6, hang=-1, main = "Dendrogram of Ward")
rect.hclust(hc_WardCereal, k=3, border = 1:3)
```
Visually examining the dendrogram, it looks as though there should be 3 clusters.





Part Three
```{r}
partitionA<- sample(1:2, nrow(Numcereal), replace = TRUE)
partitionB<- 3 - partitionA

clusterA<- cutree(hc_WardCereal, k=3)
clusterB<- cutree(hc_WardCereal, partitionB)

consistency<- sum(clusterA==clusterB)/length(clusterB)
consistency
```
There is a moderate amount of stability and clustering within this structure due to the consistency sum being 0.534.





Part Four
First, what makes a cereal "Healthy" needs to be defined. Since no definition is given, I am creating one, in which healthy cereal has fiber > 3 and protein > 3. The data should not be normalized because we are looking for specific values within the catorgies/columns to filter into the partitions. Based on this definition of a healthy cereal, only 3 cereals are considered healthy. 
```{r}
HealthyCereal<- clusterA[which(cereal$protein>3 & cereal$fiber>3)]
cat("Healthy Cereals Clustering", HealthyCereal, "/n")
```
